Gaia DR2 view of the GD-1 stellar stream

### 2018-04-28 paper/gd1.tex
abstract edits

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold streams
of stars that are extremely valuable tracers of the large- and small-scale
properties of mass around the Galaxy.
Most of such streams discovered in the Milky Way reside in its halo, and are therefore primarily sensitive to the distribution of dark matter.
% Many thin stellar streams have been discovered around the Milky Way, the
% majority of which exist in the Galactic halo and are therefore primarily
- cleanest sample, foreground removed w gaia parallaxes; members selected on combination of gaia pm and panstarrs colors
- we thus got the crispest view of the stream to date, revealing in high resolution the rich structure which was only hinted at in the previous generation of data
% We present astrometric data for one of the longest stellar streams --- the
% GD-1 stream --- from the second data release (\DR) of the \gaia\ mission.
% We re-discover significant density variations along the stream by selecting
% candidate member stars using the exquisite \gaia\ data alone, including the
% existence of a prominent, apparent gap in the stream.
% <Using photometric data from the \pans, we refine our selection ...>
These density variations --- and especially the gap --- are possible indications
of past interactions with large, massive perturbers.
<estimate the mass of interaction...>
% 
% Many thin stellar streams have been discovered around the Milky Way, the
% majority of which exist in the Galactic halo and are therefore primarily
% sensitive to the distribution of dark matter.
% Here we present astrometric data for one of the longest stellar streams --- the
% ``GD-1'' stream --- from data release 2 (\DR) of the \gaia\ mission.
% We re-discover significant density variations along the stream by selecting
% candidate member stars first using only the exquisite \gaia\ data.
% Combined with filtering using Pan-STARRS photometry, we see clear evidence of
% multiple prominent under-densities and haps in the stream.
% These density variations --- and especially the gaps --- are possible
% indications of past interactions with large, massive perturbers.
% Given that the orbit of the stream crosses the Milky Way disk at large radius,
% these perturbers are likely halo objects and would have to have masses XX,
% suggestive of dark matter sub-halos.
\end{abstract}

>>>>

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold streams
of stars that are extremely valuable tracers of the large- and small-scale
distribution of matter in the Galaxy.
Most of such streams discovered in the Milky Way reside in its halo, and are therefore primarily sensitive to the distribution of dark matter.
We present a sample of highly-probable members of the longest cold stream, GD-1, selected on small parallaxes and retrograde proper motions provided by the \gaia\ mission, and \pans\ photometry consistent with an old and metal-poor population at a distance of $\sim10\,$kpc.
This selection extended the stream by $10^\circ$(?), revealed a possible location of the progenitor and mapped significant density variations along the stream in unprecedented detail.
In addition to several prominent gaps, for the first time we also find evidence of stream members off the main track.
No(?) secular formation mechanisms predict the existence of such spurs, but they are a natural result of a stream interaction with a massive perturber.
Based on the impulse approximation arguments, we estimate that the detected underdensity and the associated spur could have been induced by a perturbation of a $\sim10^7\rm M_\odot$(?) object.
Our model of GD-1 indicates that the stream crosses the Milky Way disk at a large radius, which puts the likely origin of the perturber beyond the stellar disk.
\end{abstract}
